---
title: "p8105_hw2_ilz2105"
author: "Lulu Zhang"
date: 2019-09-29
output: github_document
---

--
__Problem 1__
--


```{r , message=FALSE}
library(tidyverse)
library(readxl)
library(tidyr)
library(dplyr)
library(ggridges)
```


### Import Mr. Trash Wheel dataset & clean sheets 

Below7
I imported the `Mr. Trash Wheel` dataset and specified the sheets: `Mr Trash Wheel`, `2017 Precipitation`, 
and `2018 Precipitation`. For the `mrtrashwheel` sheet, I omitted rows that did not contain `Dumpster` specific data 
and non-data entries. The number of `Sports Balls` was rounded to the nearest integer and converted to 
an integer variable using `as.integer`. 

Next, I read and cleaned the precipitation data for the `2018 Precipitation` and `2017 Precipitation` sheets. 
For each, I omitted rows without precipitation data and added a variable `year`. Then I combined the precipitation 
datasets and converted `month` to a character variable. 

`r median(pull(filter(mr )))

```{r}
# read and clean Mr. Trash Wheel sheet
mrtrashwheel = read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", 1, col_names = TRUE) %>% 
  janitor::clean_names() %>% 
drop_na(dumpster) %>% 
mutate(sports_balls = round(sports_balls, digits = 0)) %>%  
mutate(sports_balls = as.integer(sports_balls))
mrtrashwheel

# read and clean 2018 precipitation data
mrtrash_data_2018 = read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2018 Precipitation", skip = 1 ) %>% 
  janitor::clean_names() %>%
  drop_na() %>% 
mutate(year = as.numeric(2018))
mrtrash_data_2018

# Read and clean 2017 precipitation data
mrtrash_data_2017 = read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", na = "NA" , skip = 1) %>% 
  janitor::clean_names() %>%
  #select(-precipitation_in, x2) %>% 
  drop_na() %>% 
  mutate(year = as.numeric(2017))
mrtrash_data_2017

# combine precipitation datasets
# convert month to character variable
precipitation_data = bind_rows(mrtrash_data_2017, mrtrash_data_2018)  %>% 
  janitor::clean_names() %>% 
mutate(month = as.numeric(month),
       month = month.name[month])
precipitation_data

# or 

both_precipitation = full_join(mrtrash_data_2017, mrtrash_data_2018, 
                               by = c("month", "total", "year")) %>% 
  janitor::clean_names() %>% 
mutate(month = as.numeric(month),
       month = month.name[month])
both_precipitation
```

--
__Problem 2__
--
### Read and tidy `pols`, `snp`, and `unemployment` data and merge

First, clean the data in `pols-month.csv`. Use separate() to break up the variable mon into integer variables 
`year`, `month`, and `day`; replace `month number` with `month name`; create a `president` variable taking values 
`gop` and `dem`, and remove `prez_dem` and `prez_gop`; and remove the `day` variable.

```{r}
# import pols-month.csv and clean
pols = read.csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(month = as.numeric(month),
       month = month.name[month],
  prez_gop = recode(prez_gop, '1' = "gop" ,  '2' = "gop" ,  '0' = "dem" ),
   president = prez_gop )%>% 
  select(-prez_dem, -prez_gop, -day)

pols
```

Second, clean the data in `snp.csv` using a similar process to the above. For consistency across datasets, 
arrange according to `year` and `month`, and organize so that `year` and `month` are the leading columns.


```{r}
# import snp.csv and clean
snp = read.csv("./fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
separate(date, into = c("month","day", "year" )) %>% 
  select(starts_with("year"), everything(), -day)  %>% 
  mutate(month = month.name[as.numeric(month)]) 

snp
```

Third, I tidied the `unemployment` data so that it can be merged with the previous datasets. This process involved 
switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r}

# import and tidy unemployment data 
# switch from 'wide' to 'long' format
unemployment = read.csv("./fivethirtyeight_datasets/unemployment.csv")  %>% 
   pivot_longer(Jan:Dec, names_to = "month", values_to = "percent") %>% 
  mutate(month = match(month, month.abb),
                       month = month.name[month]) %>% 
  janitor::clean_names()

unemployment

```

Next, I joined the datasets by merging `snp` into `pols`, and merging `unemployment` into the result.

```{r}

# merge pols_month, snp, and unemployment
final = inner_join(snp, pols, by = c("year", "month")) %>% 
  janitor::clean_names() %>% 
  merge( unemployment, by = c("year", "month"))
final
```
Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).




--
__Problem 3__
--

### Import `Popular_Baby_Names` dataset, tidy, and manipulate

Load and tidy the data. Note that, although these data may seem fairly well formatted initially, 
the names of a categorical predictor and the case structure of string variables changed over time; 
you’ll need to address this in your data cleaning. Also, some rows seem duplicated, and these will 
need to be removed (hint: google something like “dplyr remove duplicate rows” to get started).


```{r}

#load baby names dataset and tidy
  #recode ethnicity and check all variables
 #name to all lowercase
baby_names = read_csv("./Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  rename("year" = year_of_birth, "name" = childs_first_name) %>% 
  mutate(
    ethnicity = recode(ethnicity, 
                     'ASIAN AND PACI' = "ASIAN AND PACIFIC ISLANDER",  
                     'BLACK NON HISP' = "BLACK NON HISPANIC", 
                     'WHITE NON HISP' = "WHITE NON HISPANIC" )) %>% 
  mutate(name = str_to_lower(name)) %>% 
  distinct() 
   

  #arrange(desc(gender))
  #recode(name = str_to_lower(name))) %>%  
  
baby_names

#rename year_of_birth to year
# rename childs_first_name to name
```
Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” 
as a female baby name over time; this should have rows for ethnicities and columns for year. Produce 
a similar table showing the most popular name among male children over time.


```{r}
#table showing rank in popularity of the name 'Olivia" 
baby_names %>% 
  #group_by(year) %>% 
  filter(name == "olivia") %>% 
  select("year", "ethnicity", "rank") %>% 
  pivot_wider(names_from = "year", values_from = "rank") %>% 
  knitr::kable(digits = 1)

# table for 'ETHAN'
baby_names %>% 
  #group_by(year) %>% 
  filter(gender == "male") %>%
  filter(rank == 1) %>% 
  select("year", "ethnicity", "rank","name" ) %>% 
  pivot_wider(names_from = "year", values_from = "name") 
  #knitr::kable(digits = 1)
```

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of 
children with a name (y axis) against the rank in popularity of that name (x axis).
```{r}

plot = baby_names %>% 
  filter( gender == "MALE", 
          ethnicity == "WHITE NON HISPANIC", 
          year == 2016) %>% 
ggplot(aes(x = rank, y = count)) +
  geom_point() 

plot
```




 

