---
title: "p8105_hw2_ilz2105"
author: "Lulu Zhang"
date: 2019-09-29
output: github_document
---

--
__Problem 1__
--


```{r}
library(tidyverse)
library(readxl)
```


## Import Mr. Trash Wheel dataset & clean sheets 

Below I imported the `Mr. Trash Wheel` dataset and specified the sheets: `Mr Trash Wheel`, `2017 Precipitation`, 
and `2018 Precipitation`. For the `mrtrashwheel` sheet, I omitted rows that did not contain `Dumpster` specific data 
and non-data entries. The number of `Sports Balls` was rounded to the nearest integer and converted to 
an integer variable using `as.integer`. 

Next, I read and cleaned the precipitation data for the `2018 Precipitation` and `2017 Precipitation` sheets. 
For each, I omitted rows without precipitation data and added a variable `year`. Then I combined the precipitation 
datasets and converted `month` to a character variable. 

```{r}
# read and clean Mr. Trash Wheel sheet
mrtrashwheel = read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", 1) %>% 
  janitor::clean_names() %>% 
drop_na(dumpster) %>% 
mutate(sports_balls = round(sports_balls, digits = 0)) %>%  
mutate(sports_balls = as.integer(sports_balls))
mrtrashwheel

# read and clean 2018 precipitation data
mrtrash_data_2018 = read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", 3 ) %>% 
  janitor::clean_names() %>%
  drop_na(x2) %>% 
mutate(year = "2018")
mrtrash_data_2018

# Read and clean 2017 precipitation data
mrtrash_data_2017 = read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", 4, na = "NA", ) %>% 
  janitor::clean_names() %>%
  drop_na( x2) %>% 
  mutate(year = "2017")
mrtrash_data_2017

# combine precipitation datasets
# convert month to character variable

```


Read and clean precipitation data for 2017 and 2018
For each, omit rows without precipitation data and add a variable year. 
Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).
```{r}

```
 

