---
title: "p8105_hw2_ilz2105"
output: github_document
---

library(tidyverse)
library(readxl)

**Problem 1**

Import Mr. Trash Wheel data & clean

```{r}
# 2018 precipitation data
mrtrash_data_2017 = read_xlsx("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", 3, na = "NA"  ) %>% 
  janitor::clean_names() %>% 
  mutate(Month = "Year")
mrtrash_data_2017

#2017 precipitation data
mrtrash_data_2018 = read_xlsx("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", 4, na = "NA") %>% 
  janitor::clean_names() %>% 
  mutate(Month = "Year")
mrtrash_data_2018

```
Read and clean the dataset:
specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
use reasonable variable names
omit rows that do not include dumpster-specific data
round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

Read and clean precipitation data for 2017 and 2018
For each, omit rows without precipitation data and add a variable year. 
Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).
```{r}

```
 

