---
title: "p8105_hw2_ilz2105"
author: "Lulu Zhang"
date: 2019-09-29
output: github_document
---

--
__Problem 1__
--


```{r}
library(tidyverse)
library(readxl)
library(tidyr)
library(dplyr)
```


## Import Mr. Trash Wheel dataset & clean sheets 

Below I imported the `Mr. Trash Wheel` dataset and specified the sheets: `Mr Trash Wheel`, `2017 Precipitation`, 
and `2018 Precipitation`. For the `mrtrashwheel` sheet, I omitted rows that did not contain `Dumpster` specific data 
and non-data entries. The number of `Sports Balls` was rounded to the nearest integer and converted to 
an integer variable using `as.integer`. 

Next, I read and cleaned the precipitation data for the `2018 Precipitation` and `2017 Precipitation` sheets. 
For each, I omitted rows without precipitation data and added a variable `year`. Then I combined the precipitation 
datasets and converted `month` to a character variable. 

```{r}
# read and clean Mr. Trash Wheel sheet
mrtrashwheel = read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", 1, col_names = TRUE) %>% 
  janitor::clean_names() %>% 
drop_na(dumpster) %>% 
mutate(sports_balls = round(sports_balls, digits = 0)) %>%  
mutate(sports_balls = as.integer(sports_balls))
mrtrashwheel

# read and clean 2018 precipitation data
mrtrash_data_2018 = read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", 3, col_names = TRUE ) %>% 
  janitor::clean_names() %>%
  drop_na(x2) %>% 
mutate(year = as.numeric(2018))
mrtrash_data_2018

# Read and clean 2017 precipitation data
mrtrash_data_2017 = read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", 4, na = "NA" , col_names = TRUE) %>% 
  janitor::clean_names() %>%
  drop_na( x2)  %>% 
  mutate(year = as.numeric(2017))
mrtrash_data_2017

# combine precipitation datasets
# convert month to character variable
precipitation_data = merge(mrtrash_data_2017, mrtrash_data_2018, by = "year")  
  #bind_rows(year, year) %>%
 # pivot_longer(
   # year_x:year_y,
    #names_to = "year" ) 
precipitation_data

precipitation_data = 
  bind_rows(mrtrash_data_2017, mrtrash_data_2018) %>%
  janitor::clean_names() %>%
  pivot_longer(
    names_to = "year" ) 
  #mutate(race = str_to_lower(race)) %>% 
  #select(movie, everything()) 
precipitation_data
```

--
__Problem 2__
--

First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables 
year, month, and day; replace month number with month name; create a president variable taking values 
gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, 
arrange according to year and month, and organize so that year and month are the leading columns.

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve 
switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key
variables take the same values.


```{r}
# import pols-month.csv and clean
pols_month = read.csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(month.name = "month_number") %>% 
  #mutate(president = c("gop", "dem")) %>% 
  drop_na(prez_dem) %>% 
  drop_na(prez_gop) %>% 
  drop_na(day)

pols_month
# import snp.csv and clean
snp = read.csv("./fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
separate(date, into = c("month","day", "year" )) %>% 
  select(starts_with("year"), everything())  %>% 
  mutate(month = month.name[as.numeric(month)])  
 #order("year", decreasing = FALSE)
snp


# import and tidy unemployment data for merge
unemployment = read.csv("./fivethirtyeight_datasets/unemployment.csv") 
unemployment
#%>% 
  janitor::clean_names() %>% 
  pivot_longer(unemployment, month, names_to = "month", names_prefix = NULL, 
               names_sep = NULL, names_pattern = NULL, names_ptypes = list(), names_repair = "check_unique", 
               values_to = "value",
               values_drop_na = FALSE, values_ptypes = list())
   #gather( month, percentage, jan:dec, factor_key = TRUE)) %>% 
  mutate(month = month.name[as.numeric(month)])
  
unemployment %>% 
order("year",na.last = FALSE, decreasing = FALSE)
unemploymentnew = unemployment[order(month),]



# merge pols_month, snp, and unemployment
```

```{r}
pols_month = read.csv("./fivethirtyeight_datasets/pols-month.csv")
pols_month
```



Join the datasets by merging snp into pols, and merging unemployment into the result.

--
__Problem 3__
--

Load and tidy the data. Note that, although these data may seem fairly well formatted initially, 
the names of a categorical predictor and the case structure of string variables changed over time; 
you’ll need to address this in your data cleaning. Also, some rows seem duplicated, and these will 
need to be removed (hint: google something like “dplyr remove duplicate rows” to get started).

Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” 
as a female baby name over time; this should have rows for ethnicities and columns for year. Produce 
a similar table showing the most popular name among male children over time.

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of 
children with a name (y axis) against the rank in popularity of that name (x axis).
```{r}

#load baby names dataset and tidy
baby_names = read_csv("./Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  distinct(baby_names,childs_first_name, .keep_all = TRUE)
baby_names
```


 

