---
title: "p8105_hw2_ilz2105"
author: "Lulu Zhang"
date: 2019-09-29
output: github_document
---

--
__Problem 1__
--


```{r , message=FALSE}
library(tidyverse)
library(readxl)
library(tidyr)
library(dplyr)
library(ggridges)
```


### Import Mr. Trash Wheel dataset & clean sheets 

Below I imported the `Mr. Trash Wheel` dataset and specified the sheets: `Mr Trash Wheel`, `2017 Precipitation`, 
and `2018 Precipitation`. For the `mrtrashwheel` sheet, I omitted rows that did not contain `Dumpster` specific data 
and non-data entries. The number of `Sports Balls` was rounded to the nearest integer and converted to 
an integer variable using `as.integer`. 

```{r, message = FALSE}
# read and clean Mr. Trash Wheel sheet
mrtrashwheel = read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", 1, col_names = TRUE) %>% 
  janitor::clean_names() %>% 
drop_na(dumpster) %>% 
mutate(sports_balls = round(sports_balls, digits = 0)) %>%  
mutate(sports_balls = as.integer(sports_balls))

```

Next, I read and cleaned the precipitation data for the `2018 Precipitation` and `2017 Precipitation` sheets. 
For each, I omitted rows without precipitation data and added a variable `year`. Then I combined the precipitation 
datasets and converted `month` to a character variable.

```{r, message = FALSE}
# read and clean 2018 precipitation data
mrtrash_data_2018 = read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2018 Precipitation", skip = 1 ) %>% 
  janitor::clean_names() %>%
  drop_na() %>% 
mutate(year = as.numeric(2018))

# Read and clean 2017 precipitation data
mrtrash_data_2017 = read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", na = "NA" , skip = 1) %>% 
  janitor::clean_names() %>%
  #select(-precipitation_in, x2) %>% 
  drop_na() %>% 
  mutate(year = as.numeric(2017))


# combine precipitation datasets
# convert month to character variable
precipitation_data = bind_rows(mrtrash_data_2017, mrtrash_data_2018)  %>% 
  janitor::clean_names() %>% 
mutate(month = as.numeric(month),
       month = month.name[month])

# or 

both_precipitation = full_join(mrtrash_data_2017, mrtrash_data_2018, 
                               by = c("month", "total", "year")) %>% 
  janitor::clean_names() %>% 
mutate(month = as.numeric(month),
       month = month.name[month])

```


The number of observations for the `mrtrashwheel` sheet was `r nrow(mrtrashwheel)`, and `r nrow(both_precipitation)` 
in the `both_precipitation` dataset. The `mrtrash_data_2018` had `r nrow(mrtrash_data_2018)` observations, and the `mrtrash_data_2017` 
had `rnrow(mrtrash_data_2017)` observations. The total precipitation in 2018 was `r sum(pull(mrtrash_data_2018, total))` inches. 
The total amount of precipitation is 2017 was `r sum(pull(mrtrash_data_2017, total))` inches.
The median number of sports balls in a dumpster in 2017 was  `r median(pull(filter(mrtrashwheel, year == "2017") , sports_balls))`.
Some key variables in the sheets are `year` and `month` because those are the variables we joined by when merging
`mrtrash_data2018` and `mrtrash_data_2018` to create `both_precipitation`.

--
__Problem 2__
--

### Read and tidy `pols`, `snp`, and `unemployment` data and merge

First, clean the data in `pols-month.csv`. Use separate() to break up the variable mon into integer variables 
`year`, `month`, and `day`; replace `month number` with `month name`; create a `president` variable taking values 
`gop` and `dem`, and remove `prez_dem` and `prez_gop`; and remove the `day` variable.

```{r, message = FALSE}
# import pols-month.csv and clean
pols = read.csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(month = as.numeric(month),
       month = month.name[month],
  prez_gop = recode(prez_gop, '1' = "gop" ,  '2' = "gop" ,  '0' = "dem" ),
   president = prez_gop
  #year = as.numeric(year)
  )%>% 
  select(-prez_dem, -prez_gop, -day)

```

Second, clean the data in `snp.csv` using a similar process to the above. For consistency across datasets, 
arrange according to `year` and `month`, and organize so that `year` and `month` are the leading columns.


```{r}
# import snp.csv and clean
snp = read.csv("./fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
separate(date, into = c("month","day", "year" )) %>% 
  select(starts_with("year"), everything(), -day)  %>% 
  mutate(month = month.name[as.numeric(month)]) 

```

Third, I tidied the `unemployment` data so that it can be merged with the previous datasets. This process involved 
switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring 
that key variables take the same values.

```{r}

# import and tidy unemployment data 
# switch from 'wide' to 'long' format
unemployment = read.csv("./fivethirtyeight_datasets/unemployment.csv")  %>% 
   pivot_longer(Jan:Dec, names_to = "month", values_to = "percent") %>% 
  mutate(month = match(month, month.abb),
                       month = month.name[month]) %>% 
  janitor::clean_names()

```

Next, I joined the datasets by merging `snp` into `pols`, and merging `unemployment` into the result.

```{r}

# merge pols, snp, and unemployment
final = inner_join(snp, pols, by = c("year", "month")) %>% 
  janitor::clean_names() %>% 
  merge( unemployment, by = c("year", "month"))

```
Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

The `pols` dataset contained information about the number of national politicians who are `dem` or `rep` at 
any given time. The data also contained info about the number of `dem` or `gop` senators, govenors, or representatives 
on the associated `date`. For example, in `march` `1947`, there are 23 republican govenors and 51 republican senators, which we see 
by looking at the `gov_gop` and `sen_gop` columns respectively for that `month` and `year`. `pols` has `r nrow(pols)` observations, with `year` 
ranging from `r range(pull(pols, year))`. 

The `snp` dataset contains `r nrow(snp)` observations with 3 variables, `year`, `month`, and `close`, which is the S&P stock
market index associated with that date. The range of years is `r range(pull(snp, year))`. 

The `unemployment` dataset contains `r nrow(unemployment)` rows and `r ncol(unemployment)` variables. The `percent` column
tells you the unemployment percentage for that associated `month` and `year`. So in June of 1948, the unemployment precentage was 3.6. 
The range of years for this data os `r range(pull(unemployment, year))`. 

Fianlly, I merged `pols`, `snp`, and `unemployment` to create `final`, which contains `r nrow(final)` observations and 
`r ncol(final)` variables. The range of years is `r range(pull(unemployment, year))` and it tells us the unemployment percentage in the `percent` column, 
S&P stock market index in the `close` column, and number of dem and rep senators/govenors/representatives
for the given `month` and `year` in the `gov_gop`, `sen_gop`, `rep_gop`, `sen_dem`, `gov_dem`, and it tells us whether the
`president` was dem or rep for the associated `year` and `month` in the `presient` column. 

--
__Problem 3__
--

### Import `Popular_Baby_Names` dataset, tidy, and manipulate

Load and tidy the data. Note that, although these data may seem fairly well formatted initially, 
the names of a categorical predictor and the case structure of string variables changed over time; 
you’ll need to address this in your data cleaning. Also, some rows seem duplicated, and these will 
need to be removed (hint: google something like “dplyr remove duplicate rows” to get started).


```{r}

#load baby names dataset and tidy
  #recode ethnicity and check all variables
 #name to all lowercase
baby_names = read_csv("./Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  rename("year" = year_of_birth, "name" = childs_first_name) %>% 
  mutate(
    ethnicity = recode(ethnicity, 
                     'ASIAN AND PACI' = "ASIAN AND PACIFIC ISLANDER",  
                     'BLACK NON HISP' = "BLACK NON HISPANIC", 
                     'WHITE NON HISP' = "WHITE NON HISPANIC" )) %>% 
  mutate(name = str_to_lower(name)) %>% 
  distinct() 

```
Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” 
as a female baby name over time; this should have rows for ethnicities and columns for year. Produce 
a similar table showing the most popular name among male children over time.


```{r}
#table showing rank in popularity of the name 'Olivia" 
olivia = baby_names %>% 
  filter(name == "olivia") %>% 
  select("year", "ethnicity", "rank") %>% 
  pivot_wider(names_from = "year", values_from = "rank") %>% 
  knitr::kable(digits = 1)
olivia
# table to find male baby name with the highest rank
popular_male_name = baby_names %>%
  filter(gender == "MALE") %>%
  filter(rank == 1) %>% 
  select(everything(), -"count") %>% 
  pivot_wider(names_from = "year", values_from = "name") %>% 
  knitr::kable(digits = 1)
popular_male_name
```
The table `olivia` shows us the frequency of the name for `year` and `ethnicity`. The rows show the `ethnicity`
and columns are `year`. 
The table `popular_male_name` shows us which male baby name was the most popular among `ethnicity` and by `year`. 
It appears that `jayden`, `ethan`, `liam`, are teh most popular as they were ranked 1 the most among the different
ethnicities and for multiple years. `jayden` was ranked 1 multiples years for `asian and pacific islander`, 
`black non hispanic`, and `hispanic`. `ethan` was also ranked 1 for multiple years among `asian and pacific islander`, 
nd `black non hispanic`.

### Scatterplot names for male, non hispanic children born in 2016

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of 
children with a name (y axis) against the rank in popularity of that name (x axis).
```{r, message = FALSE}

plot = baby_names %>% 
  filter( gender == "MALE", 
          ethnicity == "WHITE NON HISPANIC", 
          year == 2016) %>% 
ggplot(aes(x = rank, y = count, col= rank)) +
  geom_point() +
   geom_smooth(se = FALSE) +
  ggtitle("Baby name frequency vs. rank in popularity")

plot
```

The plot `Baby name frequency vs. rank in popularity` shows the number of children with a name (`count`) against the rank 
in popularity of that name (`rank`) for male, non-hispanic children born in 2016. We can see that as the rank in popularity
of that name decreases, the number of children with that name also decreases, suggesting a somewhat linear relationship. 




 

